Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections

### 2024-12-03 paper/paper.qmd
committing changes to paper.qmd in variables section

The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions. The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the Appendix.

@@

The raw data underwent pre-processing to address missing values, particularly for adjusted prices. Missing entries were filled using forward and backward imputation to ensure data continuity. Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions. The cleaned data, prepared using the `tidyverse`, `lubridate`, `TTR`, and `arrow` libraries, provides a consistent and comprehensive foundation for modeling and analysis.

Further details on the data preparation process are included in [Appendix -@sec-data-cleaning].

@@

Despite the strengths of these variables, limitations exist. Raw price points do not account for intraday variations, while constructed metrics may oversimplify complex market behaviors. External shocks, such as economic crises or regulatory changes, may not be fully captured within the variables’ scope. Acknowledging these constraints is essential for interpreting the results and understanding the broader context of the analysis. These limitations are further discussed in {@sec-data-details}.

## Variables

@@

### Outcome Variable
The **outcome variable** for this analysis is the **adjusted closing price (`adjusted`)**, which serves as the target for prediction. This represents the stock's closing price adjusted for corporate actions like stock splits or dividends, providing a normalized measure of stock value. The model aims to predict this variable based on historical data and constructed predictors.

@@

# Model

@@

# Results

@@

# Discussion

@@

# Additional data details

>>>>

The implications of these findings extend beyond price predictions, offering a framework to evaluate market behavior in response to external factors such as elections and economic conditions. The remainder of this paper is organized as follows: Section 2 describes the data sources, variables, and preprocessing steps; Section 3 outlines the modeling approach and feature engineering strategies; Section 4 presents the results and discusses trends in predicted stock prices; and Section 5 addresses the implications and limitations of the findings, with additional model diagnostics detailed in the [Appendix -@sec-model-details].

@@

The raw data underwent pre-processing to address missing values, particularly for adjusted prices. Missing entries were filled using forward and backward imputation to ensure data continuity. Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions. The cleaned data, prepared using the `tidyverse` of [@tidyverse], `lubridate` of [@lubridate], `TTR` of [@ttr], and `arrow` of [@arrow] libraries, provides a consistent and comprehensive foundation for modeling and analysis. Further details on the data preparation process are included in [Appendix -@sec-data-details].

@@

Despite the strengths of these variables, limitations exist. Raw price points do not account for intraday variations, while constructed metrics may oversimplify complex market behaviors. External shocks, such as economic crises or regulatory changes, may not be fully captured within the variables’ scope. Acknowledging these constraints is essential for interpreting the results and understanding the broader context of the analysis. These limitations are further discussed in [Appendix -@sec-data-details].

## Variables {#sec-variables}

@@

### Outcome Variable
The **outcome variable** for this analysis is the **adjusted closing price (`adjusted`)**, which serves as the target for prediction. This represents the stock's closing price adjusted for corporate actions like stock splits or dividends, providing a normalized measure of stock value. The model aims to predict this variable based on historical data and constructed predictors.
```{r}
#| echo: false
#| eval: true
#| label: fig-outcome-var
#| fig-cap: "Distribution of Adjusted Closing Prices by Stock"
#| warning: false

# Plot the distribution of the adjusted closing price, faceted by stock
ggplot(analysis_data, aes(x = adjusted)) +
  geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~symbol, scales = "free") +
  labs(
    x = "Adjusted Closing Price",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(face = "bold")
  )
```
@fig-outcome-var illustrates the distribution of the adjusted closing prices (adjusted) for each FAANG stock, faceted by stock symbol. These distributions highlight the variability in stock prices across the companies, with stocks like NFLX showing higher price ranges, whereas GOOGL and AMZN have narrower distributions. The visual emphasizes the diverse price behavior among FAANG stocks, underscoring the need for tailored models that capture these unique trends. By normalizing for corporate actions like stock splits or dividends, the adjusted closing price serves as a reliable target variable for forecasting and analysis.

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-pred-var
#| tbl-cap: "Random Sample of Predictive Variables for Modeling Adjusted Closing Price"
#| warning: false

# Select and display random sample of predictive variables from analysis_data
set.seed(123) # For reproducibility
library(dplyr)
library(knitr)

analysis_data %>%
  select(
    Lag_1,
    Rolling_Mean_7,
    sma_20,
    sma_50,
    volatility,
    daily_return,
    symbol_encoded
  ) %>%
  sample_n(5) %>% # Randomly sample 5 rows
  mutate(across(everything(), ~ round(., 2))) %>% # Round to 2 decimal places
  knitr::kable(
    col.names = c(
      "Lag_1",
      "Rolling Mean (7-Day)",
      "SMA (20-Day)",
      "SMA (50-Day)",
      "Volatility",
      "Daily Return",
      "Symbol (Encoded)"
    )
  )

```
@tbl-pred-var displays a random selection of five rows from the dataset, showcasing the predictive variables utilized in the modeling process. Each row represents data for a specific stock, including values for lagged adjusted closing prices (Lag_1), smoothed weekly averages (Rolling Mean (7-Day)), and medium- and long-term moving averages (SMA (20-Day) and SMA (50-Day)). These variables are complemented by measures of price volatility, daily percentage returns (Daily Return), and an encoded numerical identifier for the stock symbol (Symbol (Encoded)). All values are rounded to two decimal places to enhance clarity and precision. This table illustrates the diverse set of historical and technical indicators used to predict future adjusted stock prices.

# Model {#sec-model}

@@

# Results {#sec-results}

@@

# Discussion {#sec-discussion}

@@

# Additional data details {#sec-data-details}
